A user-extensible and adaptable parser architecture

### 2008-08-07 doc/sgai-2008.tex
Moved SGAI 2008 paper to this directory, to avoid replicating lots of stuff in a different directory. Made the required modifications to have it all work.

XXX REVIEWER SUGGESTED COMPARING TO EXISTING WORK, BUT I DID NOT FIND ANY
SIMILAR EXISTING WORK\@.  SHOULD SEARCH AGAIN\@.

@@

Each action contains the code required to deal with log lines from a
particular category, e.g.\ rejections.  The actions are parser-specific:
each parser author will need to write the actions required from scratch,
unless extending an existing parser.  In the Postfix parser developed for
this project there are \numberOFactions{} actions and \numberOFrules{}
rules; the distribution of rules to actions is very skewed, as shown in
\graphref{Distribution of rules per action}.  The distribution is skewed
for the same reason the architecture is so successful in enabling users to
parse new log lines: most new log lines can be parsed by a combination of a
new rule and an existing action.
\showgraph{build/plot-action-distribution}{Distribution of rules per
action}{Distribution of rules per action} Unsurprisingly the action with
the highest number of associated rules is \texttt{REJECTION}, the action
which handles Postfix rejecting a mail; it is followed by
\texttt{SAVE\_BY\_QUEUEID}, the action responsible for handling informative
inputs: these are log lines that supplement the data gathered from other
log lines (such as rejections), but that do not by themselves represent
interesting events.  The third most common action is, perhaps surprisingly,
\texttt{IGNORE}: the \texttt{IGNORE} action does nothing when executed,
allowing uninteresting log lines to be parsed without causing any effects.
The parser warns about every unparsed log line, to alert the user that they
need to alter or extend their ruleset; the Postfix parser successfully
parses every input line in the \numberOFlogFILESall{} log files it is
currently being tested with.  The remaining actions have only one or two
associated rules: in some cases this is because there will only ever be one
type of log line for that action, e.g.\ all log lines showing that a remote
client has connected are matched by a single rule and handled by the
\texttt{CONNECT} action; in other cases the action is required to address a
deficiency in the log files, or a complication which arises during parsing.

@@

An important property of a parser is how parsing time scales relative to
input size: does it scale linearly, polynomially, or exponentially?
\Graphref{Parsing time, file size, and number of lines} shows the parsing
time in seconds, file size in MB, and tens of thousands of log lines for
each of the \numberOFlogFILES{} log files.  All three lines run roughly in
parallel, giving the impression that the algorithm scales linearly with
input size.  This impression is borne out by \graphref{parsing time vs file
size vs number lines factor} which plots both the ratio of file size vs
parsing time, and the ratio of number of log lines vs parsing time (higher
is better in both cases).  The ratios are quite tightly banded across the
graph, showing that the algorithm scales linearly.  The ratio increases
(i.e.\ improves) for log files 22 and 62--68 despite their larger than
average size (shown in \graphref{Parsing time, file size, and number of
lines}); this is explained in \sectionref{Why are there dips in the
graphs?}.  \showgraph{build/plot-normal-filesize-numlines}{Parsing time,
file size, and number of lines}{Parsing time, file size, and number of
lines} \showgraph{build/plot-normal-filesize-numlines-factor}{Ratio of file
size and number of lines to parsing time}{parsing time vs file size vs
number lines factor}

@@

\Graphref{Parsing time relative to shuffled} shows the parsing times of
normal and reversed orderings relative to shuffled ordering.  Overall this
optimisation provides a modest but worthwhile reduction in parsing time of
approximately 10\%.  \showgraph{build/plot-hits}{Hits per rule}{rule hits
graph}
\showgraph{build/plot-normal-and-reverse-relative-to-shuffle}{Parsing time
relative to shuffled ordering}{Parsing time relative to shuffled}

@@

How any architecture scales as the number of rules increases is important,
but it is particularly important in this architecture because it is
expected that there will be a large number of rules.  There are
\numberOFrules{} rules in the full Postfix ruleset, whereas the minimum
number of rules required to parse the \numberOFlogFILES{} log files is
\numberOFrulesMINIMUM{}, \numberOFrulesMINIMUMpercentage{} of the total.  A
second set of statistics was generated using the minimum ruleset and the
results compared to the statistics generated using the full ruleset: the
percentage parsing time increase when using the full ruleset instead of the
minimal ruleset for normal, shuffled and reversed orderings is shown in
\graphref{Percentage parsing time increase of maximum ruleset over minimum
ruleset}.  \showgraph{build/plot-normal-vs-smaller}{Percentage parsing time
increase of maximum ruleset over minimum ruleset}{Percentage parsing time
increase of maximum ruleset over minimum ruleset}

It is clear from \graphref{Percentage parsing time increase of maximum
ruleset over minimum ruleset} that the increased number of rules has a
noticeable performance impact with reverse ordering, and to a lesser extent
with shuffled ordering.  The normal ordering (where are the most frequently
matching rules are tried first) shows a mean increase of
\input{build/normal-vs-smaller-percentage-increase.tex} in parsing time for
a \numberOFrulesMAXIMUMpercentage{} increase in the number of rules.  These
results show that the architecture scales very well as the number of rules
increases, and that sorting the rules is an important optimisation enabling
this.

@@

\bibliographystyle{spmpsci}
\bibliography{../common/bibliography}

>>>>

XXX REVIEWER SUGGESTED COMPARING TO EXISTING WORK, BUT I DID NOT FIND ANY
SIMILAR EXISTING WORK\@.  I SHOULD REALLY SEARCH AGAIN\@.

@@

Each action contains the code required to deal with log lines from a
particular category, e.g.\ rejections.  The actions are parser-specific:
each parser author will need to write the actions required from scratch,
unless extending an existing parser.  In the Postfix parser developed for
this project there are \numberOFactions{} actions and \numberOFrules{}
rules; the distribution of rules to actions is very skewed, as shown in
\graphref{Distribution of rules per action}.  The distribution is skewed
for the same reason the architecture is so successful in enabling users to
parse new log lines: most new log lines can be parsed by a combination of a
new rule and an existing action.
\showgraph{build/graph-action-distribution}{Distribution of rules per
action}{Distribution of rules per action} Unsurprisingly the action with
the highest number of associated rules is \texttt{REJECTION}, the action
which handles Postfix rejecting a mail; it is followed by
\texttt{SAVE\_BY\_QUEUEID}, the action responsible for handling informative
inputs: these are log lines that supplement the data gathered from other
log lines (such as rejections), but that do not by themselves represent
interesting events.  The third most common action is, perhaps surprisingly,
\texttt{IGNORE}: the \texttt{IGNORE} action does nothing when executed,
allowing uninteresting log lines to be parsed without causing any effects.
The parser warns about every unparsed log line, to alert the user that they
need to alter or extend their ruleset; the Postfix parser successfully
parses every input line in the \numberOFlogFILESall{} log files it is
currently being tested with.  The remaining actions have only one or two
associated rules: in some cases this is because there will only ever be one
type of log line for that action, e.g.\ all log lines showing that a remote
client has connected are matched by a single rule and handled by the
\texttt{CONNECT} action; in other cases the action is required to address a
deficiency in the log files, or a complication which arises during parsing.

@@

An important property of a parser is how parsing time scales relative to
input size: does it scale linearly, polynomially, or exponentially?
\Graphref{Parsing time, file size, and number of lines} shows the parsing
time in seconds, file size in MB, and tens of thousands of log lines for
each of the \numberOFlogFILES{} log files.  All three lines run roughly in
parallel, giving the impression that the algorithm scales linearly with
input size.  This impression is borne out by \graphref{parsing time vs file
size vs number lines factor} which plots both the ratio of file size vs
parsing time, and the ratio of number of log lines vs parsing time (higher
is better in both cases).  The ratios are quite tightly banded across the
graph, showing that the algorithm scales linearly.  The ratio increases
(i.e.\ improves) for log files 22 and 62--68 despite their larger than
average size (shown in \graphref{Parsing time, file size, and number of
lines}); this is explained in \sectionref{Why are there dips in the
graphs?}.  \showgraph{build/graph-input-size-vs-parsing-time}{Parsing time,
file size, and number of lines}{Parsing time, file size, and number of
lines} \showgraph{build/graph-input-size-vs-parsing-time-ratio}{Ratio of
file size and number of lines to parsing time}{parsing time vs file size vs
number lines factor}

@@

\Graphref{Parsing time relative to shuffled} shows the parsing times of
normal and reversed orderings relative to shuffled ordering.  Overall this
optimisation provides a modest but worthwhile reduction in parsing time of
approximately 10\%.  \showgraph{build/graph-hits}{Hits per rule}{rule hits
graph} \showgraph{build/graph-optimal-and-reverse-vs-shuffle}{Parsing time
relative to shuffled ordering}{Parsing time relative to shuffled}

\newpage{}

@@

How any architecture scales as the number of rules increases is important,
but it is particularly important in this architecture because it is
expected that there will be a large number of rules.  There are
\numberOFrules{} rules in the full Postfix ruleset, whereas the minimum
number of rules required to parse the \numberOFlogFILES{} log files is
\numberOFrulesMINIMUM{}, \numberOFrulesMINIMUMpercentage{} of the total.  A
second set of statistics was generated using the minimum ruleset and the
results compared to the statistics generated using the full ruleset: the
percentage parsing time increase when using the full ruleset instead of the
minimal ruleset for normal, shuffled and reversed orderings is shown in
\graphref{Percentage parsing time increase of maximum ruleset over minimum
ruleset}.
\showgraph{build/graph-full-ruleset-vs-minimum-ruleset}{Percentage parsing
time increase of maximum ruleset over minimum ruleset}{Percentage parsing
time increase of maximum ruleset over minimum ruleset}

It is clear from \graphref{Percentage parsing time increase of maximum
ruleset over minimum ruleset} that the increased number of rules has a
noticeable performance impact with reverse ordering, and to a lesser extent
with shuffled ordering.  The normal ordering (where are the most frequently
matching rules are tried first) shows a mean increase of
\input{build/include-full-ruleset-vs-minimum-ruleset} in parsing time for a
\numberOFrulesMAXIMUMpercentage{} increase in the number of rules.  These
results show that the architecture scales very well as the number of rules
increases, and that sorting the rules is an important optimisation enabling
this.

@@

\bibliographystyle{spmpsci}
\bibliography{logparser-bibliography.bib}
